---
title: "Clean 'Traps' data"
author: "indra deCastro"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    theme: united
---

<style type="text/css">
.table {
    width: 100%;
}
</style>

```{r setup, include=FALSE}
# setwd("/home/indra/Documents/20160101.Noriega.exp.remocion/")
knitr::opts_knit$set(root.dir ="/home/indra/Documents/20160101.Noriega.exp.remocion/", message=F, warning=F, verbose=F, comment=F)
knitr::opts_chunk$set(message=F, warning=F, verbose=F, echo=F, autodep=T, cache=T, fig.width=10)
options(markdown.HTML.header = system.file("misc", "datatables.html", package = "knitr"))
```

## Import data
```{r import, echo=T}
raw <- read.table("data/traps.csv", header=T, sep="\t", check.names=F, na.strings=c("","-"))
```

## Basic editing
```{r, results=F}
raw$manejo <- factor(raw$manejo, labels=c("Int", "Ext"))
raw$no_trap <- as.factor(raw$no_trap)

# change all NA values into ZEROS
raw[is.na(raw)] <- 0
names(raw)[1:5]
# cat(paste(sum(raw[-c(1:3)]), "individuals in the whole table"))

# check that no names have spaces
# grep(" ", names(raw), value=T, perl=F)

# reorder columns so that "site" is the first column
# NOT USED as it adds numbers at the end of colnames to ensure uniqueness
# raw <- raw[,c("site", names(raw)[names(raw)!="site"])]
```

## Study species (du/tri/...)-plicates
```{r sp.dup, results=F}
# change BIG "S" to lowercase "s" in species names with "sp"
# names(raw)[grep("Sp[0-9]$", names(raw))] <- gsub("_S", "_s", grep("Sp[0-9]$", names(raw), value=T), fixed=T)

# species that appear in more than one column
# and have given a name with "sp[0-9]"
spsp <- sort(grep("sp[0-9]$", names(raw), value=T))
dupspsp <- spsp[duplicated(spsp)]
dupspcols <- unique(unlist(lapply(dupspsp, function(x){which(names(raw)==x)})))

# add site name at end of species
names(raw)[dupspcols] <- sapply(dupspcols, function(sp){
      spsites <- raw$codigo_parcela[which(raw[,sp]!=0)]
      spsites <- unique(strtrim(as.character(spsites), 5))
      paste0(names(raw)[sp], ".", spsites)
})

# check that none of the species with "sp" is duplicated
any(duplicated(sort(grep("sp[0-9]$", names(raw), value=T)))==T)
```

## Check for typos in species names
```{r}
# select names that DO NOT contain "sp"
nosp <- grep("sp[0-9]", names(raw)[-c(1:3)], value=T, invert=TRUE)

# fuzzy searching of duplicates
possible.typos <- lapply(unique(nosp), function(x){agrep(paste0("^", x, "$"), unique(nosp), value=T, max.distance=0.2)})
```

## Merge columns with same species name
```{r, results=F}
aggr <- as.data.frame(do.call(cbind, by(t(raw[-c(1:3)]), INDICES=names(raw)[-c(1:3)], FUN=colSums)))
# cat(paste(sum(aggr), "individuals in the whole table"))
small <- cbind("site"=strtrim(raw$codigo_parcela, 5), raw[1:3], aggr)
names(small)[1:5]
# cat(paste(sum(small[-c(1:4)]), "individuals in the whole table"))
```

## Save clean and checked table for future use


# Data summary
```{r, echo=F, results=F}
abund.sp <- apply(small[-c(1:4)], 2, sum)
range(abund.sp)
summary(abund.sp)
```

## General
Number of sites: `r length(unique(small$site))`  
Number of traps: `r length(unique(paste0(small$site, small$no_trap)))`  
Number of species: `r length(names(small))`  
Total abundance: `r as.integer(sum(small[-c(1:4)]))`  

## Abundance by species
```{r}
topsp <- data.frame("abund"=sort(apply(small[-c(1:4)], 2, sum), decreasing=T))
library(DT)
datatable(topsp, rownames=T)
```

### Species abundance
```{r}
summary(abund.sp)
```

### Number of singletons: `r length(which(apply(small[-c(1:4)], 2, sum)==1))`

### Most widely sampled species
```{r}
trap.sp <- data.frame("number of traps present"=sort(apply(small[-c(1:4)], 2, function(x){length(which(x!=0))}), decreasing=T))
datatable(trap.sp, rownames=T)
```

## By site
### Total abundance by site and treatment
```{r}
abund.trap <- cbind(small[1:3], "abund"=apply(small[-c(1:4)], 1, sum))
library(plyr)
abund.sites <- ddply(abund.trap, .(codigo_parcela), summarize,
                     "tot"=sum(abund),
                     "mean"=mean(abund),
                     "sd"=sd(abund))

tot.abund.site <- data.frame("site" = unique(abund.trap$site),
      "Int" = abund.sites[grep("Int", abund.sites$codigo_parcela),"tot"],
      "Ext" = abund.sites[grep("Ext", abund.sites$codigo_parcela),"tot"])

datatable(tot.abund.site)
```

### Summary for abundance by treatment
```{r}
# round(rbind("min"=apply(tot.abund.site[,c("Int", "Ext")], 2, min),
#       "mean"=apply(tot.abund.site[,c("Int", "Ext")], 2, mean),
#       "sd"=apply(tot.abund.site[,c("Int", "Ext")], 2, sd),
#       "max"=apply(tot.abund.site[,c("Int", "Ext")], 2, max),
#       "tot"=apply(tot.abund.site[,c("Int", "Ext")], 2, sum)),2)

abund.summary <- round(rbind(
      cbind("Int"=c(summary(tot.abund.site$Int)[1:4], "Stdev."=sd(tot.abund.site$Int), summary(tot.abund.site$Int)[5:6]), 
            "Ext"=c(summary(tot.abund.site$Ext)[1:4], "Stdev."=sd(tot.abund.site$Ext), summary(tot.abund.site$Ext)[5:6])),
      "Tot."=apply(tot.abund.site[,c("Int", "Ext")], 2, sum)
      ),2)

datatable(abund.summary)
```


```{r writedata}
names(small)[1:4] <- c("site", "code", "manejo", "trap")

## Write new clean table
write.csv(small, paste0("data/", format(Sys.time(), '%Y%m%d'), ".traps.csv"), row.names=F, quote=F)
```




