---
title: "Clean 'Traits' data"
author: "indra deCastro"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    theme: united
---

```{r setup, include=FALSE}
# setwd("/home/indra/Documents/20160101.Noriega.exp.remocion/")
knitr::opts_knit$set(root.dir ="/home/indra/Documents/20160101.Noriega.exp.remocion/", message=F, warning=F, verbose=F, comment=F)
knitr::opts_chunk$set(message=F, warning=F, verbose=F, echo=F, autodep=T, cache=T, fig.width=10)
# options(markdown.HTML.header = system.file("misc", "datatables.html", package = "knitr"))
```

## Import data
```{r import, echo=T}
traits <- read.table("data/traits.csv", header=T, sep=",", na.strings="-")
traps <- read.table("data/traps.csv", header=T, sep=",")
```

## Mini summary
Number of sp in traps: `r length(traps[,-c(1:4)])`  
Number of sp in traits: `r length(names(table(traits$sp)))`  

## Traits without species
```{r}
# names(traps)[270:278]; names(traps)[1:5]
sptraps <- names(traps[,-c(1:4)])
sptraits <- names(table(traits$sp))
setdiff(sptraits, sptraps)
# traitless <- setdiff(sptraps, sptraits)
# traitless

sitexsp <- aggregate(traps[-c(1:4)], by=list(traps$site), sum)
# tlesssites <- sitexsp[,traitless]
# rownames(tlesssites) <- levels(traps$site)
# t(tlesssites[apply(tlesssites, 1, sum)!=0,])

# (we have no measurements for 5sp)
# (or 7sp, counting Bolbelasmus and Trox)
# sps <- nrow(table(traits$sp))
# nrow(sps)
# # names of those 60 species and numb of measures per species
# sps[order(sps$freq),]
# table(sps[order(sps$freq),"freq"])
```

## Species without traits
```{r}
setdiff(sptraps, sptraits)
```

```{r}
##### very important chunk #####
# used further down
library(plyr)
avg <- ddply(traits, .(trait, sp, code), summarize,
             "n" = length(value),
             "avg" = mean(value),
             "sd" = sd(value))
```


## Species without traits in some sites
```{r}
# aggregate(traps[,-c(1:4)], list(traps$code), FUN=function(x){apply(x, 2, sum)})
etiq.cols <- 1:4
praw <- traps[,-etiq.cols]
etiq <- traps[,etiq.cols]
abun <- aggregate(praw, by=list(traps$code), sum)
library(reshape)
labun <- melt(abun)
labun <- labun[labun$value!=0,]
names(labun) <- c("code", "sp", "abun")


# nrows should match, right??
uavg <- unique(avg[,c("sp", "code")])
# head(labun)
# head(uavg)
# 
# dim(uavg)
# dim(labun)

pabun <- paste(labun$sp,labun$code)
pavg <- paste(uavg$sp,uavg$code)

arrange(labun[!pabun %in% pavg,], code, sp)
```


## Species and site

- There are many species with 10 measurements. These are singletons (1 individual * 10 measurements).  
- There are many species (193) with 100 measurements. This should be the max if Noriegas protocol was correctly followed (10 individuals no matter which treatment * 10 measures).  

BUT...  
...some researchers didn't understand and measured **more** than 10 individuals per species. 

- In these cases, maximum can be 10 individuals in Int and 10 individuals in Ext.  
Thus, 20 individuals measured 10 traits = 200 measures / per sp / per site.  
```{r}
sps <- data.frame(table(traits$sp, substr(traits$code, 1, 5)))
sps <- sps[sps$Freq!=0,]
ss <- t(t(table(sps[order(sps$Freq),"Freq"])))
ss <- data.frame("measures per site"=rownames(ss), "number of sp"=ss[,1], row.names=NULL)
library(DT)
datatable(ss, rownames=F, options = list(pageLength = 20, dom="t"), width=15)
# maximum can be 10 individuals in Int and 10 individuals in Ext
# thus, 20 individuals measured 10 traits = 200 measures / per sp / per site
```

## Measurements per species
```{r}
sps <- data.frame(table(traits$sp))
sps <- sps[sps$Freq!=0,]
# datatable(sps[order(sps$Freq, decreasing=T),], rownames=F, colnames=c("Species", "Number of measurements"), escape=FALSE, options = list(searching = FALSE))

datatable(sps[order(sps$Freq, decreasing=T),], colnames=c("Species", "Number of measurements"), rownames=F, escape=FALSE, options = list(searching=F))
```

## Measurements per species and trait
```{r}
# sort(apply(traps[,-c(1:4)], 2, sum))

# table(traits[traits$sp=="Digitonthophagus_gazella","trait"])
# spxtrait <- aggregate(traits$value, list("sp"=traits$sp), function(x){table(x[complete.cases(x)])})
myt <- traits[complete.cases(traits),]
spxtrait <- aggregate(myt$trait, list("sp"=myt$sp), table)
spxtrait <- cbind("sp"=spxtrait$sp, data.frame(spxtrait$x))
datatable(arrange(spxtrait, desc(spxtrait$el)), rownames=F)
```

## Some trait values are missing
```{r}
options(htmlwidgets.TOJSON_ARGS = list(na = 'string'))
datatable(traits[is.na(traits$value),c(1,4,5,7,8)], option=list(dom="t"), rownames=F)
```






```{r, eval=F}
# Other mistakes that were solved
absp <- paste0(labun$sp, ".", labun$code)
trsp <- paste0(uavg$sp, ".", uavg$code)
# setdiff(absp, trsp)
setdiff(trsp, absp)

# ejemplo
traps[traps$code=="USA01Int",c("code", "Alloblackburneus_rubeolus"), drop=F]
traits[traits$code=="USA01Int" & traits$sp=="Alloblackburneus_rubeolus",c(1,4,7,8)]
grep("rubeolus", trsp, value=T)
grep("rubeolus", absp, value=T)

# pero otras 4-5 comprobaciones tb se cumplen
# Canthidium_pseudopuncticolle.MEX03Int
# Onthophagus_similis.SPA01Int
# Aphodius_fimetarius.SPA02Ext
# Onthophagus_buculus.BRA02Int
```

```{r, eval=F}
mex05 <- traps[traps$site=="MEX05",]
mex05.named <- t(mex05[,c(1:4, 4+which(apply(mex05[,5:length(mex05)],2,sum)!=0))])
mex05 <- mex05[,4+which(apply(mex05[,5:length(mex05)],2,sum)!=0)]
sum(mex05)
t(mex05)
apply(mex05, 2, sum)


mex06 <- traps[traps$site=="MEX06",]
mex06.named <- t(mex06[,c(1:4, 4+which(apply(mex06[,5:length(mex06)],2,sum)!=0))])
mex06 <- mex06[,4+which(apply(mex06[,5:length(mex06)],2,sum)!=0)]
sum(mex06)
t(mex06)
apply(mex06, 2, sum)
```
