---
title: "Clean 'Removal' data"
author: "indra deCastro"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    theme: united
---

<style type="text/css">
.table {
    width: 40%;
}
</style>

```{r setup, include=FALSE}
# setwd("/home/indra/Documents/20160101.Noriega.exp.remocion/")
knitr::opts_knit$set(root.dir ="/home/indra/Documents/20160101.Noriega.exp.remocion/", message=F, warning=F, verbose=F, comment=F)
knitr::opts_chunk$set(message=F, warning=F, verbose=F, echo=F, autodep=T, cache=T, fig.width=10)
options(markdown.HTML.header = system.file("misc", "datatables.html", package = "knitr"))
```

## Import data
```{r import, echo=T}
raw <- read.table("data/removal.csv", header=T, sep=",", na.strings=c("-", "N/A", "NA"))
```

## Check for table consistency
i.e. check for: Number of Sites? number of treatments? number of experimental units?...  
Number of sites: `r length(unique(raw$codigo_parcela))/2`
```{r labels}
labels <- 1:4
apply(raw[labels], 2, table)
```
Why are treatments called "1/2" instead of "Int/Ext"??  
I suggest that anything that is NOT a number (=continous variables) should be given a name (with letters).  

```{r renametreatments, echo=T}
raw$manejo <- factor(raw$manejo, labels=c("Int", "Ext"))
table(raw$manejo)
```


## Explore data
i.e. check for: ranges of data make sense? no evident outliers? number of NA values?...
```{r exploredata}
summary(raw[-labels])
```

Summarize percentage of NAs in each variable:
```{r}
t(t(data.frame("perc.NAs"=apply(raw[-labels], 2, function(x){round((nrow(raw)-length(which(complete.cases(x)))) / nrow(raw) *100, 2)}))))
```

### All variables
```{r othervars}
summary(raw[-labels])

## put histograms on the diagonal
panel.hist <- function(x, ...)
{
    usr <- par("usr"); on.exit(par(usr))
    par(usr = c(usr[1:2], 0, 1.5) )
    h <- hist(x, plot = FALSE)
    breaks <- h$breaks; nB <- length(breaks)
    y <- h$counts; y <- y/max(y)
    rect(breaks[-nB], 0, breaks[-1], y, col = "cyan", ...)
}
pairs(raw[-labels], bg = "grey80", diag.panel = panel.hist, cex.labels = 1, upper.panel=NULL)
# pairs(raw[-labels])
```

<!-- ### dung_final_dry -->
<!-- There is an obvious outlier in data point in *dung_final_dry*:   -->
```{r outliers, eval=F}
plot(raw$final_dung_dry, pch=c(1,19)[as.factor(raw$final_dung_dry==max(raw$final_dung_dry, na.rm=T))], col=c("black", "red")[as.factor(raw$final_dung_dry==max(raw$final_dung_dry, na.rm=T))], cex=c(1,2)[as.factor(raw$final_dung_dry==max(raw$final_dung_dry, na.rm=T))])

t(raw[which(raw$final_dung_dry==max(raw$final_dung_dry, na.rm=T)),])
```
<!-- Is is a typo when digitalizing data?   -->
<!-- Look at the values in the replicates of this experiment:   -->
```{r, eval=F}
plot(raw[raw$codigo_parcela=="AUS02Int","final_dung_dry"])
summary(raw[raw$codigo_parcela=="AUS02Int","final_dung_dry"])

# correct 
raw[which(raw$final_dung_dry==max(raw$final_dung_dry, na.rm=T)), "final_dung_dry"] <- 161
```
<!-- Should the first "1" be deleted? Thus, the correct value be 161.0?   -->
<!-- I have corrected it for future exploration, but NOT in the original data table.   -->

### Dry variables
Moreover,  
The variable "dry_paper_bag"  
`r raw[which(raw$dry_paper_bag>20),"dry_paper_bag"]`  
and "cont_dung_dry" have a problem:  
`r raw[which(raw$cont_dung_dry>200),"cont_dung_dry"]`  
Most probably it is due to dragging the initial value (25.89 and 207.13) in Excel to complete the other values.  
However, this 207.13gr seems an outlier for that variable.  
**Is this possible?** Might it be 20.713 ??

```{r testdrypaper}
raw$newdrypaper <- raw$dry_paper_bag
raw$newdrypaper[raw$newdrypaper>25.89] <- 25.89

raw$newcontdung <- raw$cont_dung_dry
raw$newcontdung[raw$newcontdung>=207.13] <- 20.713

pairs(raw[c("initial_dung", "final_dung_wet", "final_dung_dry", "newcontdung", "paper_bag", "newdrypaper")], bg = "grey80", diag.panel = panel.hist, cex.labels = 1, upper.panel=NULL)
```

### Variables of interest
```{r}
clean.removal <- raw[,c("codigo_parcela", "manejo", "unidad", "no_unidad", "initial_dung", "final_dung_wet", "final_dung_dry")]
clean.removal$manejo <- substring(clean.removal$codigo_parcela, 6,9)
names(clean.removal) <- c("site", "manejo", "type", "unit", "initial_dung_wet", "final_dung_wet", "final_dung_dry")
# write.table(clean.removal, "data/clean.removal.csv", row.names=F, quote=F, sep=";")


raw <- raw[-grep(x=names(raw), pattern=c("dry|bag|new"))]
```

### Initial dung
```{r initialdung}
hist(raw$initial_dung, main="Weight of initial dung (grams)", xlab="grams of wet dung", ylab="number of samples", breaks=seq(min(raw$initial_dung, na.rm=T)-1, max(raw$initial_dung, na.rm=T)+1, 0.5)); box()

table(cut(raw$initial_dung, breaks=c(100,299.9,300,500), labels=c("smaller", "300gr", "bigger")))
```

Most of the samples (`r round(length(which(raw$initial_dung==300.0)) / nrow(raw) *100, 2)`%) weigh exactly 300.0 grams.  
However, there are `r length(which(raw$initial_dung>300.0))` samples (`r round(length(which(raw$initial_dung>300.0)) / nrow(raw) *100, 2)`%) over that weight.

### Final dung (wet)
```{r}
hist(raw$final_dung_wet, main="Weight of final dung (grams)", xlab="grams of wet dung", ylab="number of samples", breaks=seq(min(raw$final_dung_wet, na.rm=T)-1, max(raw$final_dung_wet, na.rm=T)+1, 2)); box()

# table(cut(raw$final_dung_wet, breaks=c(0,299.9,300,320), labels=c("smaller", "300gr", "bigger")))
```

## METHOD 1
The complete removal formula is:  
*Removal rate = (ExpIniWet - (ExpFinWet + WaterLoss)) / 48h*  
Where *WaterLoss* is calculated as *ConIniWet - ConFinWet*  using the control experiments.  

### Water loss (evaporation in control experiments)
```{r realremoval, fig.height=20, fig.width=15}
library(plyr)
waterdif <- ddply(raw, .(codigo_parcela, unidad, manejo, no_unidad), .fun=summarize, 
                  "waterdif" = initial_dung - final_dung_wet)

waterdif.cont <- waterdif[waterdif$unidad=="Cont",]
waterdif.cont$ID <- strtrim(waterdif.cont$codigo_parcela, width=5)

# significance
water.pval <- ddply(waterdif.cont, .(ID), summarize,
                    "pval"=wilcox.test(waterdif ~ manejo, exact=F)$p.value)
                    

waterdif.cont <- merge(waterdif.cont, water.pval)
waterdif.cont <- arrange(waterdif.cont, pval)
waterdif.cont$ID <- factor(waterdif.cont$ID, levels=unique(waterdif.cont$ID))

# plotting
library(ggplot2)
library(ggsignif)
white.violins <- ggplot(waterdif.cont) +
      aes(x=manejo, y=waterdif) +
      geom_violin() +
      facet_wrap(~ID, ncol=6) +
      labs(x="manejo", y="water loss", title="Evaporation in each treatment") +
      scale_y_continuous(expand = c(0,0), limits=c(0,400)) +
      geom_signif(comparisons = list(c("Int", "Ext")), map_signif_level=TRUE)

white.violins
```
Note that the water loss is different (wilcoxon test) between **`r round(100*nrow(water.pval[which(water.pval$pval<0.05),])/nrow(water.pval),2)`%** Intensive and Extensive treatments.  
**Is this reasonable??**  
Are there important habitat differences between Intensive and Extensive treatments?  
**Where Control experiments really beetle-free??**



### Removal rate
Are these observations of simple dung difference correct?  
```{r dungaddition}
xp <- raw[raw$unidad!="Cont",]
dungdif <- xp$initial_dung - xp$final_dung_wet
plot(dungdif, pch=c(1,19)[as.factor(dungdif<0)], col=c("black", "red")[as.factor(dungdif<0)]); abline(h=0)

qplot(dungdif, geom = "histogram", fill = dungdif > 0, binwidth=4, xlab="weight diff (water not considered)")
```
Dung increased during experiment, there was **NO** removal, but addition of dung!  
Percentage of experiments with weight increase: `r round(100*nrow(xp[which(dungdif<0),])/nrow(xp),2)`%  

```{r}
library(DT)
xp[which(dungdif<0),1:6] %>% 
      datatable(rownames=F, elementId="five")
```
In Canadian experiments the difference might be due to method inaccuracy (soil crumbs not well brushed enough etc.)

**This percentage is even higher if we consider water loss:**
```{r removalhist, echo=F}
H2O.loss <- ddply(waterdif[waterdif$unidad=="Cont",], .(codigo_parcela, manejo), summarise,
                  "H2O.loss" = mean(waterdif, na.rm=T))
xp <- merge(xp, H2O.loss[,c("codigo_parcela", "H2O.loss")])
xp <- xp[,c("codigo_parcela", "manejo", "unidad", "no_unidad", "initial_dung", "final_dung_wet", "H2O.loss")]

xp$rem48 <- (xp$initial_dung - (xp$final_dung_wet + xp$H2O.loss))/48

qplot(rem48, data=xp, geom = "histogram", fill = rem48 > 0, binwidth=0.1, xlab="Removal rate in 48h (gr/h)")
```
Percentage of experiments with weight increase: `r round(100*nrow(xp[which(xp$rem48<0),])/nrow(xp),2)`%  

Table **only showing experiments with negative** removal rate (=*increase of weight*)
```{r}
library(DT)
xp[which(xp$rem48<0),] %>%
      datatable(rownames=F, elementId="sixtyone") %>%
      formatRound(columns=c("H2O.loss", "rem48"), digits=2)

xp$ID <- strtrim(xp$codigo_parcela, width=5)
```



## METHOD 2
*dry.rem48 = (initial_dung_dry – final_dung_dry) / 48h*  
*initial_dung_dry =  initial_dung_wet \* (1 – water_proportion)*  
*water_proportion = (final_dung_wet – final_dung_dry) / final_dung_wet*
```{r, eval=F}
met2 <- raw[,c("codigo_parcela", "manejo", "unidad", "initial_dung", "final_dung_wet", "final_dung_dry")]
```


## Export data
```{r, eval=F}
rem.site <- ddply(xp, .(codigo_parcela, ID, manejo), summarise,
                  "avg.wet.rem48" = mean(rem48),
                  "std.wet.rem48" = sd(rem48))

rem.site$manejo <- substring(rem.site$codigo_parcela, 6, 9)
head(rem.site)
names(rem.site) <- c(c("code", "site"), names(rem.site)[3:5])
all <- Reduce(f=merge, x=list(abun.site, rich.site, even.site, rem.site))
head(all)
# write.table(all, "data/all.csv", row.names=F, quote=F, sep=";")
```



## Removal METHOD 1 according to treatment
```{r removal.plot, fig.height=20, fig.width=15}
# usexp <- xp[which(xp$rem48>0),]
usexp <- xp


# significance
usexp.pval <- ddply(usexp, .(ID), summarize,
                    "pval"=wilcox.test(rem48 ~ manejo, exact=F)$p.value)
# arrange(usexp.pval, pval)
                    
usexp <- merge(usexp, usexp.pval)
usexp <- arrange(usexp, pval)

sig.countries <- c("WAL03", "IRA01", "SPA01", "SPA02", "WAL02", "MEX03", "ITA01", "MEX05", "MEX04", "COL03", "CHI01", "POR01", "FRA01", "ARG01")
my.ord <- c(sig.countries, setdiff(unique(usexp$ID), sig.countries))
usexp$ID <- factor(usexp$ID, levels=my.ord, ordered=T)

# plotting
library(ggplot2)
library(ggsignif)

white.violins <- usexp %>% 
      ggplot() +
      aes(x=manejo, y=rem48, fill=manejo) +
      geom_violin() +
      # geom_boxplot() + 
      facet_wrap(~ID, ncol=6) +
      labs(x="manejo", y="Removal rate 48h (gr/h)", title="Dung Removal in each treatment") +
      scale_y_continuous(expand = c(0,0), limits=c(0,7.5)) +
      geom_signif(comparisons = list(c("Int", "Ext")), map_signif_level=T, test="wilcox.test", test.args=list(exact=F))

white.violins
# arrange(usexp.pval, pval)
```
Note that the removal rate calculated in 48h is ONLY different (two-sided unpaired wilcoxon test) **`r nrow(usexp.pval[which(usexp.pval$pval<0.05),])`/`r nrow(usexp.pval)`** **(`r round(100*nrow(usexp.pval[which(usexp.pval$pval<0.05),])/nrow(usexp.pval),2)`%)** between Intensive and Extensive treatments.  
**Is this reasonable??**  
If no other statistical tool provides differences, it is useless searching environmental explanatory variables for such *homogeneous response variable*.


```{r, eval=F}
usexp <- xp


# significance
usexp.pval <- ddply(usexp, .(ID), summarize,
                    "pval"=wilcox.test(rem48 ~ manejo, exact=F)$p.value)
arrange(usexp.pval, pval)
                    
usexp <- merge(usexp, usexp.pval)
usexp <- arrange(usexp, pval)

sig.countries <- c("WAL03", "IRA01", "SPA01", "SPA02", "WAL02", "MEX03", "ITA01", "MEX05", "MEX04", "COL03", "CHI01", "POR01", "FRA01", "ARG01")
my.ord <- c(sig.countries, setdiff(unique(usexp$ID), sig.countries))
usexp$ID <- factor(usexp$ID, levels=my.ord, ordered=T)

# my try
mywx <- function(x,y){
      tmp <- wilcox.test(x, y, exact=F, paired=F)
      tmp$p.value <- p.adjust(tmp$p.value, n = 2, method = "BH")
      tmp
}  

# plotting
library(ggplot2)
library(ggsignif)

white.violins <- usexp %>% 
      ggplot() +
      aes(x=manejo, y=rem48, fill=manejo) +
      geom_violin() +
      # geom_boxplot() + 
      facet_wrap(~ID, ncol=6) +
      labs(x="manejo", y="Removal rate 48h (gr/h)", title="Dung Removal in each treatment") +
      scale_y_continuous(expand = c(0,0), limits=c(0,7.5)) +
      geom_signif(comparisons = list(c("Int", "Ext")), map_signif_level=T, test="mywx")

white.violins

```


<!-- ## Paired Wilcoxon test -->
<!-- **reports even less significant differences in removal between treatments -->
```{r paired.wilcox, eval=F}
wwil <- reshape(usexp, direction="wide", drop=names(usexp)[-c(1,3,5,9)],
        timevar="manejo", idvar=c("ID", "no_unidad"))

# significance
library(plyr)
usexp.pval.paired <- ddply(wwil, .(ID), summarize,
                    "pval.paired"=wilcox.test(x=rem48.Ext, y=rem48.Int, paired=T, exact=F)$p.value)

usexp.pval.paired <- arrange(usexp.pval.paired, pval.paired)
bothpvals <- merge(usexp.pval.paired, usexp.pval)
arrange(bothpvals, pval.paired)
```

<!-- trying to plot paired removal rates -->
```{r, fig.height=20, fig.width=15, eval=F}
library(ggpubr)
p <- ggpaired(usexp, x="manejo", y="rem48", group=1,
          color = "manejo", palette = "jco", add="jitter",
          line.color = "gray", line.size = 0.4,
          facet.by = "ID", short.panel.labs = FALSE)
# Use only p.format as label. Remove method name.
p + stat_compare_means(method = "wilcox.test", label = "p.format", paired = TRUE, method.args=list(exact="F"))
```

