---
title: "Clean 'Sites' data"
author: "indra deCastro"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
      html_document:
      toc: true
theme: united
---
      
```{r setup, include=FALSE}
# setwd("/home/indra/Documents/20160101.Noriega.exp.remocion/")
knitr::opts_knit$set(root.dir ="/home/indra/Documents/20160101.Noriega.exp.remocion/", message=F, warning=F, verbose=F, comment=F)
knitr::opts_chunk$set(message=F, warning=F, verbose=F, echo=F, autodep=T, cache=T, fig.width=10)
# options(markdown.HTML.header = system.file("misc", "datatables.html", package = "knitr"))
```

## Import data
```{r import, echo=T, eval=T}
raw <- read.table("data/sites.csv", header=T, sep=",", dec=",", na.strings="?", stringsAsFactors=F)
```

### Names of all variables
- BIO1 = Annual Mean Temperature
- BIO2 = Mean Diurnal Range (Mean of monthly (max temp - min temp))
- BIO3 = Isothermality (- BIO2/- BIO7) (* 100)
- BIO4 = Temperature Seasonality (standard deviation *100)
- BIO5 = Max Temperature of Warmest Month
- BIO6 = Min Temperature of Coldest Month
- BIO7 = Temperature Annual Range (- BIO5-- BIO6)
- BIO8 = Mean Temperature of Wettest Quarter
- BIO9 = Mean Temperature of Driest Quarter
- BIO10 = Mean Temperature of Warmest Quarter
- BIO11 = Mean Temperature of Coldest Quarter
- BIO12 = Annual Precipitation
- BIO13 = Precipitation of Wettest Month
- BIO14 = Precipitation of Driest Month
- BIO15 = Precipitation Seasonality (Coefficient of Variation)
- BIO16 = Precipitation of Wettest Quarter
- BIO17 = Precipitation of Driest Quarter
- BIO18 = Precipitation of Warmest Quarter
- BIO19 = Precipitation of Coldest Quarter

### 
```{r, eval=F}
library(raster)
library(sp)

extract.point.values <- function(coords, var="bio"){
      site <- coords[1]
      coords <- coords[2:3]
      
      r <- getData("worldclim", var=var, res=0.5, lon=as.numeric(coords[1]), lat=as.numeric(coords[2]), path="/media/indra/KREAS/my20180315temp/")
      
      # OPTIONAL show data at current moment
      plot(r[[1]], main=site)
      points <- SpatialPoints(data.frame(t(as.numeric(coords))), proj4string = r@crs)
      points(points, pch="+")
      text(coords, site, cex=0.8, pos=4)
      
      # name layers and extract values in our sites
      names(r) <- paste("BIO", 1:19, sep=".")
      points <- SpatialPoints(data.frame(t(as.numeric(coords))), proj4string = r@crs)
      values <- extract(r, points)
      rm(r)
      
      # save data 
      values <- data.frame(values, row.names=site)
      return(values)
}

wclim2df <- function(coords){
      sites <- coords$site
      print(as.character(sites))
      if(names(coords)[2]!="x" & names(coords)[3]!="y"){
            cat("Provide coordinates in a dataframe with names 'x' (=longitude) and 'y' (=latitude)")
      }
      lall <- apply(coords, 1, function(x){extract.point.values(x)})
      cbind(sites, do.call(rbind, lall))
}
```

### Download and Extract data for all sites
```{r}
coords <- as.data.frame(apply(raw[,c("lon", "lat")], 2, as.numeric))
coords <- cbind(raw$site, coords)
names(coords) <- c("site", "x", "y"); head(coords)
n <- 1:nrow(coords)
# n <- 75:nrow(coords) # solo para pruebas
coords <- coords[n,]

# NOT RUN !! interent intensive!!
# run only once to download from web
# myclimate <- wclim2df(coords)
```

### Export all climatic values to an excel sheet
```{r}
# write.table(myclimate, "data/bioclim.csv", row.names=F, quote=F, sep=";")
```




















## Plot all sites
```{r, results="asis", cache=F}
# needed for googleVis::gvisGeoChart
sites <- data.frame("LatLong"=paste(raw$lat, raw$lon, sep=":"))
sites$tip <- raw$site
sites$tip <- paste(paste(raw$site),
                   paste("Alt:", raw$alt, "m"),
                   # paste("Prec:", raw$prec, "mm/año"),
                   # paste("Tmean:", raw$tavg, "ºC"),
                   # paste("Raref.Rich:", div$raref.rich, "species"),
                   # paste("Beta.tot:", round(div$mpair.Btot, 2)),
                   # paste("Bray.tot:", round(div$mpair.BrayTot, 2)),
                   sep="<BR>")


library(googleVis)
library(knitr)
op <- options(gvis.plot.tag='chart')
# op <- options(gvis.print.tag='chart')
intmap <- gvisMap(sites[,], "LatLong" , "tip", 
                  options=list(showTip=TRUE, 
                               showLine=TRUE, 
                               enableScrollWheel=TRUE,
                               mapType='terrain', 
                               useMapTypeControl=TRUE))
plot(intmap)
```

***
      ***