---
title: "Correlations"
author: "indra deCastro"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    theme: united
---
      
<style type="text/css">
      .table {
            width: 100%;
      }
</style>
      
```{r setup, include=FALSE}
# setwd("/home/indra/Documents/20160101.Noriega.exp.remocion/")
knitr::opts_knit$set(root.dir ="/home/indra/Documents/20160101.Noriega.exp.remocion/", message=F, warning=F, verbose=F, comment=F)
knitr::opts_chunk$set(message=F, warning=F, verbose=F, echo=F, autodep=T, cache=T, fig.width=10)
options(markdown.HTML.header = system.file("misc", "datatables.html", package = "knitr"))
```

## Import data
```{r import, echo=T}
raw <- read.table("data/all.csv", header=T, sep=";")
```

```{r, echo=F}
foc.names <- c("abun", "rich", "even", # taxonomic
               paste0("BIO.", 1:19), # climatic
               "rem.met1", "rem.met2", # removal
               "lat", "altitud", # geographical
               "land.history", "cow.dens", "anthelmintic", # manegement
               "Fric.behav", "Feve.behav", "Fdis.behav", "RaoQ.behav", "mpd.behav", # functional for 9 relocationXsize categories
               # "Fdiv.behav" cant be calculated if only categorical traits are used 
               "Fric.morph", "Feve.morph", "Fdiv.morph", "Fdis.morph", "RaoQ.morph", "mpd.morph") # functional for morphological traits
foc <- raw[,foc.names]
# dim(foc)
```

## Select variables to explore correlations
Variables of interest (`r ncol(foc)`):

- taxonomic: abun, rich, even
- climatic: `r paste0("BIO.", 1:19)` 
- removal: rem.met1, rem.met2
- geographical: lat, altitud
- manegement: land.history, cow.dens, anthelmintic
- functional for 9 relocationXsize categories: Fric.behav, Feve.behav, Fdis.behav, RaoQ.behav, mpd.behav  
(*Fdiv.behav cant be calculated if only categorical traits are used*)
- functional for morphological traits: Fric.morph, Feve.morph, Fdiv.morph, Fdis.morph, RaoQ.morph, mpd.morph

Names of environmental variables:  

- BIO1 = Annual Mean Temperature
- BIO2 = Mean Diurnal Range (Mean of monthly (max temp - min temp))
- BIO3 = Isothermality (BIO2/BIO7) (* 100)
- BIO4 = Temperature Seasonality (standard deviation *100)
- BIO5 = Max Temperature of Warmest Month
- BIO6 = Min Temperature of Coldest Month
- BIO7 = Temperature Annual Range (BIO5--BIO6)
- BIO8 = Mean Temperature of Wettest Quarter
- BIO9 = Mean Temperature of Driest Quarter
- BIO10 = Mean Temperature of Warmest Quarter
- BIO11 = Mean Temperature of Coldest Quarter
- BIO12 = Annual Precipitation
- BIO13 = Precipitation of Wettest Month
- BIO14 = Precipitation of Driest Month
- BIO15 = Precipitation Seasonality (Coefficient of Variation)
- BIO16 = Precipitation of Wettest Quarter
- BIO17 = Precipitation of Driest Quarter
- BIO18 = Precipitation of Warmest Quarter
- BIO19 = Precipitation of Coldest Quarter

## Calculate correlations
```{r, echo=F}
library(dplyr)
library(reshape2)

allcor <- as.matrix(cor(foc, use="pairwise.complete.obs"))
allcor.melt <- arrange(melt(allcor), -abs(value))
# allcor.melt <- melt(allcor)
allcor.uni <- allcor.melt[seq(1,nrow(allcor.melt), 2),]
```

## Filter correlations
```{r}
filcor <- dplyr::filter(allcor.uni,
              Var1 != Var2 & # identity corelations
              paste0(substr(Var1, 1,3), substr(Var2, 1,3)) != "BIOBIO" &
              paste0(substr(Var1, 1,1), substr(Var2, 1,1)) != "FF"
              )
```

### Highest correlations
Criteria:  

- 0.5 as arbitrary threshold
- correlation between climatic variables (both climatic)
```{r}
filcor0.5 <- dplyr::filter(filcor, 
                           value > 0.5 |
                           value < -0.5) # .5 is arbitrary

library(DT)
datatable(filcor0.5, rownames=F, options = list(pageLength = nrow(filcor0.5), dom="t")) %>% formatRound(columns=c("value"), digits=3)
```

### All correlations for removal
*Correlation between removal and all other variables regardless of the low correlation value*
```{r}
removal.cor <- dplyr::filter(filcor,
              substr(Var1, 1,7) == "rem.met" |
              substr(Var2, 1,7) == "rem.met")
rem.left <- removal.cor[grep("rem.met", removal.cor$Var1),]
rem.right <- removal.cor[grep("rem.met", removal.cor$Var2),]
rem.right <- with(rem.right, data.frame("Var1"=Var2, "Var2"=Var1, value))
arr.rem.cor <- arrange(rbind(rem.left, rem.right), Var1, Var2)

datatable(arr.rem.cor, rownames=F) %>% formatRound(columns=c("value"), digits=3)
```

