---
title: "Paired comparisons Wilcoxon and t-Test"
author: "indra deCastro"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
      toc: true
      theme: united
---
      
<style type="text/css">
      .table {
            width: 100%;
      }
</style>
      
```{r setup, include=FALSE}
# setwd("/home/indra/Documents/20160101.Noriega.exp.remocion/")
knitr::opts_knit$set(root.dir ="/home/indra/Documents/20160101.Noriega.exp.remocion/", message=F, warning=F, verbose=F, comment=F)
knitr::opts_chunk$set(message=F, warning=F, verbose=F, echo=F, autodep=T, cache=T, fig.width=10)
options(markdown.HTML.header = system.file("misc", "datatables.html", package = "knitr"))
```

## Import data
```{r import, echo=T}
raw <- read.table("data/all.csv", header=T, sep=";")
```

## Select variables to compare
```{r, echo=F}
foc.names <- c("abun", "rich", "even", # taxonomic
               # paste0("BIO.", 1:19), # climatic
               "rem.met1", "rem.met2", # removal
               "altitud", # geographical
               "land.history", "cow.dens", "anthelmintic", # manegement
               "Fric.behav", "Feve.behav", "Fdis.behav", "RaoQ.behav", "mpd.behav", # functional for 9 relocationXsize categories
               # "Fdiv.behav" cant be calculated if only categorical traits are used 
               "Fric.morph", "Feve.morph", "Fdiv.morph", "Fdis.morph", "RaoQ.morph", "mpd.morph") # functional for morphological traits
# foc <- raw[,foc.names]
# dim(foc)
```

Variables of interest (`r length(foc.names)`):

- taxonomic: abun, rich, even
<!-- - climatic: `r paste0("BIO.", 1:19)`  -->
- removal: rem.met1, rem.met2
<!-- - geographical: lat, altitud -->
- manegement: land.history, cow.dens, anthelmintic
- functional for 9 relocationXsize categories: Fric.behav, Feve.behav, Fdis.behav, RaoQ.behav, mpd.behav  
(*Fdiv.behav cant be calculated if only categorical traits are used*)
- functional for morphological traits: Fric.morph, Feve.morph, Fdiv.morph, Fdis.morph, RaoQ.morph, mpd.morph

<!-- Names of environmental variables:   -->

<!-- - BIO1 = Annual Mean Temperature -->
<!-- - BIO2 = Mean Diurnal Range (Mean of monthly (max temp - min temp)) -->
<!-- - BIO3 = Isothermality (BIO2/BIO7) (* 100) -->
<!-- - BIO4 = Temperature Seasonality (standard deviation *100) -->
<!-- - BIO5 = Max Temperature of Warmest Month -->
<!-- - BIO6 = Min Temperature of Coldest Month -->
<!-- - BIO7 = Temperature Annual Range (BIO5--BIO6) -->
<!-- - BIO8 = Mean Temperature of Wettest Quarter -->
<!-- - BIO9 = Mean Temperature of Driest Quarter -->
<!-- - BIO10 = Mean Temperature of Warmest Quarter -->
<!-- - BIO11 = Mean Temperature of Coldest Quarter -->
<!-- - BIO12 = Annual Precipitation -->
<!-- - BIO13 = Precipitation of Wettest Month -->
<!-- - BIO14 = Precipitation of Driest Month -->
<!-- - BIO15 = Precipitation Seasonality (Coefficient of Variation) -->
<!-- - BIO16 = Precipitation of Wettest Quarter -->
<!-- - BIO17 = Precipitation of Driest Quarter -->
<!-- - BIO18 = Precipitation of Warmest Quarter -->
<!-- - BIO19 = Precipitation of Coldest Quarter -->

## Natural log +1 transform all the variables of interest
```{r}
raw[,foc.names] <- apply(raw[,foc.names], 2, function(x){log(x+1)})
```


## Calculate paired Wilcoxon test
```{r}
# Example for one variable with NA
# tmp <- raw[,c("site", "manejo", "even")]
# tmp <- tmp[complete.cases(tmp),]
# tmp <- raw[rep(table(tmp$site)==2, each=2),c("manejo", "even")]
# with(tmp, wilcox.test(even~manejo, paired=T, exact=F, na.action=)$p.value)

all.wilcox <- sapply(foc.names, function(var){
      
      tmp <- complete.cases(raw[,c("manejo", var)])
      tmp <- raw[,c("site", "manejo", var)]
      tmp <- tmp[complete.cases(tmp),]
      tmp <- raw[rep(table(tmp$site)==2, each=2),c("manejo", var)]
      wilcox.test(tmp[,var] ~ tmp$manejo, paired=T, exact=F)$p.value})

all.wilcox <- data.frame("site"=names(all.wilcox), "pval"=all.wilcox, row.names=NULL)
# library(dplyr)
# ord.all.wilcox <- arrange(all.wilcox, pval)
```


```{r, eval=F}
library(DT)
datatable(ord.all.wilcox, width=35, options = list(pageLength = nrow(ord.all.wilcox), dom="t")) %>%
      formatRound(columns=c("pval"), digits=4)
```

## Calculate NOT paired Wilcoxon test
```{r}
unpaired.wilcox <- sapply(foc.names, function(var){
      tmp <- complete.cases(raw[,c("manejo", var)])
      tmp <- raw[,c("site", "manejo", var)]
      tmp <- tmp[complete.cases(tmp),]
      tmp <- raw[rep(table(tmp$site)==2, each=2),c("manejo", var)]
      wilcox.test(tmp[,var] ~ tmp$manejo, paired=F, exact=F)$p.value})

unpaired.wilcox <- data.frame("site"=names(unpaired.wilcox), "unpaired.wilcox"=unpaired.wilcox, row.names=NULL)
```



## Calculate paired t-Test
```{r}
all.ttest <- sapply(foc.names, function(var){
      tmp <- complete.cases(raw[,c("manejo", var)])
      tmp <- raw[,c("site", "manejo", var)]
      tmp <- tmp[complete.cases(tmp),]
      tmp <- raw[rep(table(tmp$site)==2, each=2),c("manejo", var)]
      t.test(tmp[,var] ~ tmp$manejo, paired=T, exact=F)$p.value})

all.ttest <- data.frame("site"=names(all.ttest), "pval"=all.ttest, row.names=NULL)
# library(dplyr)
# ord.all.ttest <- arrange(all.ttest, pval)
```

## Calculate NOT paired t-Test
```{r}
unpaired.ttest <- sapply(foc.names, function(var){
      tmp <- complete.cases(raw[,c("manejo", var)])
      tmp <- raw[,c("site", "manejo", var)]
      tmp <- tmp[complete.cases(tmp),]
      tmp <- raw[rep(table(tmp$site)==2, each=2),c("manejo", var)]
      t.test(tmp[,var] ~ tmp$manejo, paired=F, exact=F)$p.value})

unpaired.ttest <- data.frame("site"=names(unpaired.ttest), "unpaired.ttest"=unpaired.ttest, row.names=NULL)
```



```{r}
# Reduce(function(x, y) merge(x, y, by="Species"), list(iris, iris, iris))
both <- Reduce(function(x, y){
      merge(x, y, by="site", suffixes=c(".wilcox", ".ttest", ".unpaired.wilcox", ".unpaired.ttest"))},
      list(all.wilcox, all.ttest, unpaired.wilcox, unpaired.ttest))
# Reduce(merge, list(all.wilcox, all.ttest, unpaired.wilcox, unpaired.ttest, by="site", suffixes=c(".wilcox", ".ttest", ".unpaired.wilcox", ".unpaired.ttest")))
# both <- merge(all.wilcox, all.ttest, by="site", suffixes=c(".wilcox", ".ttest"))
# names(both) <- c("variable", "pval.wilcox", "pval.ttest")

library(dplyr)
library(DT)
datatable(arrange(both, pval.wilcox), width=35, options = list(pageLength = nrow(all.ttest), dom="t")) %>%
      formatRound(columns=c("pval.wilcox", "pval.ttest", "unpaired.wilcox", "unpaired.ttest"), digits=4)
```

## Paired plots
```{r, eval=F}
numvars <- 1:length(foc.names)
long <- melt(raw[c("manejo", foc.names[numvars])])
long <- arrange(long, variable)
both <- arrange(both, variable)

mylabs <- levels(long$variable)
levels(long$variable) <- paste0(mylabs[numvars],
                               " (w=", round(both$pval.wilcox[numvars], 2),
                               ", t=", round(both$pval.ttest[numvars], 2), ")")

ggplot(long, aes_(x=~manejo, y=~value)) +
      # geom_boxplot() +
      geom_violin() +
      # geom_line(aes(group = manejo)) + 
      facet_wrap(~variable, scales="free") +
      geom_dotplot(binaxis='y', stackdir='center', dotsize=0.5) +
      # geom_point() +
      labs(x="", y="", title="")

```



```{r pairedplot, eval=T}
library(reshape2)
library(PairedData)

plot.paired <- function(var) {
  tmp <- raw[,c("site", "manejo", var)]
  # although normally dcast is used to summarize df, here I use it as reshape to wide format
  exvar <- dcast(tmp, site ~ manejo)
  pd <- with(log10(exvar[,c("Int", "Ext")]), paired(Ext, Int))
  plot(pd, type = "profile") +
        labs(x="", y=var, title=toupper(var))
        # + theme_bw()
}

par(mfrow=c(4,5), new=T)
lapply(foc.names, plot.paired)
```


```{r, eval=F}
long <- melt(raw[2:6])
exvar <- reshape(long, direction="wide", timevar="manejo", idvar=c("site", "variable"), v.names="value")
names(exvar) <- c("site", "var", "Ext", "Int")
exvar <- exvar[complete.cases(exvar),]

with(exvar, plot(paired(Ext, Int), groups=var))

library(ggpubr)
ggpaired(exvar, cond1=exvar$Ext, cond2=exvar$Int, facet.by="var", fill = "condition", palette = "jco")

before <-c(200.1, 190.9, 192.7, 213, 241.4, 19.69, 172.2, 185.5, 205.2, 193.7)
after <-c(392.9, 393.2, 345.1, 393, 434, 42.79, 422, 383.9, 392.3, 352.2)
groups <- rep(letters[1:5], 2)

d <- data.frame("before" = before, "after" = after, "groups" = groups)
# ggpaired(d, cond1 = "before", cond2 = "after", fill = "condition", palette = "jco")
ggpaired(d, cond1 = "before", cond2 = "after",
    fill = "condition", palette = "jco", facet.by="groups")

```




```{r, eval=F}
library(ggplot2)
# library(ggsignif)

white.violins <- raw[,c("manejo", "abun")] %>% 
      ggplot() +
      aes(x=manejo, y=log10(abun)) +
      geom_violin() +
      # geom_boxplot() + 
      # facet_wrap(~ID, ncol=6) +
      labs(x="manejo", y="Beetle abundance by site", title="PAIRED comparison") +
      # scale_y_continuous(expand = c(0,0), limits=c(0,7.5)) +
      # geom_signif(comparisons = list(c("Int", "Ext")), map_signif_level=F, test="wilcox.test", test.args=list(exact=F, paired=T))

white.violins + geom_dotplot(binaxis='y', stackdir='center', dotsize=0.5)
```




