---
title: "Removal methods"
author: "indra deCastro"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    theme: united
---
      
<style type="text/css">
      .table {
            width: 40%;
      }
</style>
      
```{r setup, include=FALSE}
# setwd("/home/indra/Documents/20160101.Noriega.exp.remocion/")
knitr::opts_knit$set(root.dir ="/home/indra/Documents/20160101.Noriega.exp.remocion/", message=F, warning=F, verbose=F, comment=F)
knitr::opts_chunk$set(message=F, warning=F, verbose=F, echo=F, autodep=T, cache=T, fig.width=10)
options(markdown.HTML.header = system.file("misc", "datatables.html", package = "knitr"))
```

## Import data
```{r import, echo=T}
raw <- read.table("data/clean.removal.csv", header=T, sep="\t", na.strings=c("-", "N/A", "NA"))
```

## METHOD 1

- wet.rem48 = (initial_dung_wet - final_dung_wet - evap.rain.orgs) / 48h
- evap.rain.orgs = initial_dung_wet - final_dung_wet **(using only CONROL units)**
```{r, echo=T}
# calculate waterloss
library(plyr)
evap.rain.orgs <- ddply(raw, .(site, type, manejo, unit), .fun=summarize, 
                  "evap.rain.orgs" = initial_dung_wet - final_dung_wet)

# only use loss of weight of control units (=theoretically waterloss)
evap.rain.orgs.controls <- evap.rain.orgs[evap.rain.orgs$type=="Cont",]
avg.evap.rain.orgs <- ddply(evap.rain.orgs.controls, .(site, manejo), summarise,
                  "avg.evap.rain.orgs" = mean(evap.rain.orgs, na.rm=T))

# merge waterloss with dung data
xp <- merge(raw, avg.evap.rain.orgs)

# calculate wet.rem48
xp$wet.rem48 <- with(xp, (initial_dung_wet - final_dung_wet - avg.evap.rain.orgs) / 48)
```

## METHOD 2
- dry.rem48 = (initial_dung_dry – final_dung_dry) / 48h
- initial_dung_dry =  initial_dung_wet * (1 – final_water_proportion)
- final_water_proportion = (final_dung_wet – final_dung_dry) / final_dung_wet  

*assumes that water proportion is the same at the end and beginning of experiment*
```{r, echo=T}
# calculate water.proportion
xp$final.water.prop <- with(xp, (final_dung_wet - final_dung_dry) / final_dung_wet)
xp$initial_dung_dry <- with(xp, initial_dung_wet * (1 - final.water.prop)) # assumes that water proportion is the same at the end and beginning of experiment
xp$dry.rem48 <- with(xp, (initial_dung_dry - final_dung_dry) / 48 )
```

## Export results at experimental-unit level
```{r}
write.table(xp, "data/unit.removal.csv", row.names=F, quote=F, sep=";")
```


## Calculate Removal by Dung Beetle and Other Organism
```{r, echo=T}
xp$unit <- as.factor(xp$unit)

rem.site <- ddply(xp, .(site, manejo, type), numcolwise(mean, na.rm=T))

all <- rem.site[rem.site$type=="Exp",]
org <- rem.site[rem.site$type=="Cont",]
```

```{r}
intvars <- c("site", "manejo", "wet.rem48", "dry.rem48")
all.rem <- merge(all[,intvars], org[,intvars], by=c("site", "manejo"), suffixes=c(".all", ".nonBeetle"))
names(all.rem)[1] <- "code"
all.rem$site <- substr(all.rem$code, 1, 5)
all.rem <- all.rem[,c("code", "site", "manejo", "wet.rem48.all", "wet.rem48.nonBeetle", "dry.rem48.all", "dry.rem48.nonBeetle")]

all.rem$dry.rem48.DungBeetles <- with(all.rem, dry.rem48.all - dry.rem48.nonBeetle)
```

<!-- ## METHOD 3 -->
<!-- (combination of METHOD 1 & 2)   -->

<!-- - dry2.rem48 = (initial_dung_wet - final_dung_wet - total_nonbeetle_loss) / 48h -->
<!-- - total_nonbeetle_loss = initial_dung_wet – final_dung_dry **(using only CONROL units)** -->

```{r}

```


## Export results at site level
```{r}
all.rem$wet.rem48.Evap.Orgs <- round(as.numeric(format(all.rem$wet.rem48.Evap.Orgs, scientific = FALSE)), 4)
write.table(all.rem, "data/site.removal.csv", row.names=F, quote=F, sep=";")
```





```{r}
xp$unit <- as.factor(xp$unit)

all <- xp[xp$type=="Exp",]
org <- xp[xp$type=="Cont",]
org.avg <- ddply(org, .(site, manejo, type), numcolwise(mean, na.rm=T))

intvars <- c("site", "manejo", "wet.rem48", "dry.rem48")
all.rem <- merge(all[,intvars], org.avg[,intvars], by=c("site", "manejo"), suffixes=c(".all", ".nonBeetle"))
names(all.rem)[1] <- "code"
all.rem$site <- substr(all.rem$code, 1, 5)
all.rem <- all.rem[,c("code", "site", "manejo", "wet.rem48.all", "wet.rem48.nonBeetle", "dry.rem48.all", "dry.rem48.nonBeetle")]

all.rem$dry.rem48.DungBeetles <- with(all.rem, dry.rem48.all - dry.rem48.nonBeetle)

# rem.site <- ddply(xp, .(site, manejo, type), numcolwise(mean, na.rm=T))
```

## Plot Removal Method 1 (wet removal)
```{r, eval=F, fig.height=20, fig.width=15}
sig.countries <- c("WAL03", "IRA01", "SPA01", "SPA02", "WAL02", "MEX03", "COL03", "ITA01", "CHI01", "POR01", "MEX01", "MEX02", "MEX06", "SAL01", "MEX05", "MEX04", "FRA01", "AUS02", "BRA03", "CAN01")
my.ord <- c(sig.countries, setdiff(unique(all.rem$site), sig.countries))
all.rem$site <- factor(all.rem$site, levels=my.ord, ordered=T)

library(dplyr)
library(ggplot2)
library(ggsignif)
white.violins <- all.rem %>% 
      ggplot() +
      aes(x=manejo, y=wet.rem48.all, fill=manejo) +
      geom_violin() +
      # geom_boxplot() + 
      facet_wrap(~site, ncol=6) +
      labs(x="manejo", y="Removal method 1 (wet removal)", title="UNPAIRED comparison") +
# WARNING: setting ylim in ggplot returns an error when data is beyond those limits
      scale_y_continuous(expand = c(0,0), limits=c(-3,9)) +
# WARNING: setting ylim in ggplot returns an error when data is beyond those limits
      geom_signif(comparisons = list(c("Int", "Ext")), map_signif_level=T, test="wilcox.test", test.args=list(exact=F, paired=F))

white.violins + geom_dotplot(binaxis='y', stackdir='center', dotsize=0.5)
```

<!-- ## Paired comparison -->
```{r, eval=F}
rem.site <- ddply(all.rem, .(code, manejo), numcolwise(mean, na.rm=T))

library(dplyr)
white.violins <- rem.site %>% 
      ggplot() +
      aes(x=manejo, y=wet.rem48.all, fill=manejo) +
      geom_violin() +
      # geom_boxplot() + 
      # facet_wrap(~site, ncol=6) +
      labs(x="manejo", y="Removal method 1 (wet removal)", title="PAIRED comparison") +
# WARNING: setting ylim in ggplot returns an error when data is beyond those limits
      scale_y_continuous(expand = c(0,0), limits=c(-3,7)) +
# WARNING: setting ylim in ggplot returns an error when data is beyond those limits
      geom_signif(comparisons = list(c("Int", "Ext")), map_signif_level=T, test="wilcox.test", test.args=list(exact=F, paired=T))

white.violins + geom_dotplot(binaxis='y', stackdir='center', dotsize=0.5)
```


## Plot Removal Method 2 (dry removal)
```{r, eval=T, fig.height=20, fig.width=15}
sig.countries <- c("WAL03", "IRA01", "SPA01", "CHI01", "FRA01", "SAL01",
                   "WAL02", "COL03", "POR01", "MEX01",  "MEX02",  "MEX06", "MEX04", "AUS02", "AUS01", "HUN01", "IND01",
                   "MEX03", "ITA01", "MEX05", "BRA03", "CAN01", "SPA02", "BRA01", "COL01", "ECU01", "WAL01")
my.ord <- c(sig.countries, setdiff(unique(all.rem$site), sig.countries))
all.rem$site <- factor(all.rem$site, levels=my.ord, ordered=T)

white.violins <- all.rem %>% 
      ggplot() +
      aes(x=manejo, y=dry.rem48.DungBeetles, fill=manejo) +
      geom_violin() +
      # geom_boxplot() + 
      facet_wrap(~site, ncol=6) +
      labs(x="manejo", y="Removal method 2 (dry removal)", title="UNPAIRED comparison") +
# WARNING: setting ylim in ggplot returns an error when data is beyond those limits
      scale_y_continuous(expand = c(0,0), limits=c(-1.5,5.5)) +
# WARNING: setting ylim in ggplot returns an error when data is beyond those limits
      geom_signif(comparisons = list(c("Int", "Ext")), map_signif_level=T, test="wilcox.test", test.args=list(exact=F, paired=F))

white.violins + geom_dotplot(binaxis='y', stackdir='center', dotsize=0.5)
```

```{r, eval=T}
all.rem.pval <- ddply(all.rem, .(site), summarize,
                    "pval"=wilcox.test(dry.rem48.DungBeetles ~ manejo, exact=F, paired=F)$p.value,
                    "W"=wilcox.test(dry.rem48.DungBeetles ~ manejo, exact=F, paired=F)$statistic)
ord.rem.pval <- arrange(all.rem.pval, pval)

library(DT)
datatable(ord.rem.pval, width=35, options = list(pageLength = nrow(ord.rem.pval), dom="t")) %>%
      formatRound(columns=c("pval"), digits=4)
```

<!-- ## Paired comparison -->
```{r}
library(dplyr)
library(ggplot2)
library(ggsignif)

rem.site <- ddply(all.rem, .(code, manejo), numcolwise(mean, na.rm=T))

white.violins <- rem.site %>% 
      ggplot() +
      aes(x=manejo, y=dry.rem48.DungBeetles) +
      geom_violin() +
      # geom_boxplot() + 
      # facet_wrap(~site, ncol=6) +
      labs(x="manejo", y="Removal method 2 (dry removal)", title="PAIRED comparison") +
# WARNING: setting ylim in ggplot returns an error when data is beyond those limits
      scale_y_continuous(expand = c(0,0), limits=c(-1,4)) +
# WARNING: setting ylim in ggplot returns an error when data is beyond those limits
      geom_signif(comparisons = list(c("Int", "Ext")), map_signif_level=F, test="wilcox.test", test.args=list(exact=F, paired=T))

white.violins + geom_dotplot(binaxis='y', stackdir='center', dotsize=0.5)
```

