---
title: "Removal methods"
author: "indra deCastro"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    theme: united
---
      
<style type="text/css">
      .table {
            width: 40%;
      }
</style>
      
```{r setup, include=FALSE}
# setwd("/home/indra/Documents/20160101.Noriega.exp.remocion/")
knitr::opts_knit$set(root.dir ="/home/indra/Documents/20160101.Noriega.exp.remocion/", message=F, warning=F, verbose=F, comment=F)
knitr::opts_chunk$set(message=F, warning=F, verbose=F, echo=F, autodep=T, cache=T, fig.width=10)
options(markdown.HTML.header = system.file("misc", "datatables.html", package = "knitr"))
```

## Import data
```{r import, echo=T}
raw <- read.table("data/clean.removal.csv", header=T, sep=";", na.strings=c("-", "N/A", "NA"))
```

## METHOD 1

- wet.rem48 = (initial_dung_wet - final_dung_wet - evap.rain.orgs) / 48h
- evap.rain.orgs = initial_dung_wet - final_dung_wet **(using only CONROL units)**
```{r, echo=T}
# calculate waterloss
library(plyr)
evap.rain.orgs <- ddply(raw, .(site, type, manejo, unit), .fun=summarize, 
                  "evap.rain.orgs" = initial_dung_wet - final_dung_wet)

# only use loss of weight of control units (=theoretically waterloss)
evap.rain.orgs.controls <- evap.rain.orgs[evap.rain.orgs$type=="Cont",]
avg.evap.rain.orgs <- ddply(evap.rain.orgs.controls, .(site, manejo), summarise,
                  "avg.evap.rain.orgs" = mean(evap.rain.orgs, na.rm=T))

# merge waterloss with dung data
xp <- merge(raw, avg.evap.rain.orgs)

# calculate wet.rem48
xp$wet.rem48 <- with(xp, (initial_dung_wet - final_dung_wet - avg.evap.rain.orgs) / 48)
```

## METHOD 2
- dry.rem48 = (initial_dung_dry – final_dung_dry) / 48h
- initial_dung_dry =  initial_dung_wet * (1 – final_water_proportion)
- final_water_proportion = (final_dung_wet – final_dung_dry) / final_dung_wet  

*assumes that water proportion is the same at the end and beginning of experiment*
```{r, echo=T}
# calculate water.proportion
xp$final.water.prop <- with(xp, (final_dung_wet - final_dung_dry) / final_dung_wet)
xp$initial_dung_dry <- with(xp, initial_dung_wet * (1 - final.water.prop)) # assumes that water proportion is the same at the end and beginning of experiment
xp$dry.rem48 <- with(xp, (initial_dung_dry - final_dung_dry) / 48 )
```

## Export results at experimental-unit level
```{r}
write.table(xp, "data/unit.removal.csv", row.names=F, quote=F, sep=";")
```


## Calculate Removal by Dung Beetle and Other Organism
```{r, echo=T}
xp$unit <- as.factor(xp$unit)
rem.site <- ddply(xp, .(site, manejo, type), numcolwise(mean, na.rm=T))

all <- rem.site[rem.site$type=="Exp",]
org <- rem.site[rem.site$type=="Cont",]
```

```{r}
intvars <- c("site", "manejo", "wet.rem48", "dry.rem48")
all.rem <- merge(all[,intvars], org[,intvars], by=c("site", "manejo"), suffixes=c(".all", ".nonBeetle"))
names(all.rem)[1] <- "code"
all.rem$site <- substr(all.rem$code, 1, 5)
all.rem <- all.rem[,c("code", "site", "manejo", "wet.rem48.all", "wet.rem48.nonBeetle", "dry.rem48.all", "dry.rem48.nonBeetle")]

all.rem$dry.rem48.DungBeetles <- with(all.rem, dry.rem48.all - dry.rem48.nonBeetle)
```

## METHOD 3
(combination of METHOD 1 & 2)  

- dry2.rem48 = (initial_dung_wet - final_dung_wet - total_nonbeetle_loss) / 48h
- total_nonbeetle_loss = initial_dung_wet – final_dung_dry **(using only CONROL units)**

```{r}

```




## Export results at site level
```{r}
all.rem$wet.rem48.Evap.Orgs <- round(as.numeric(format(all.rem$wet.rem48.Evap.Orgs, scientific = FALSE)), 4)
write.table(all.rem, "data/site.removal.csv", row.names=F, quote=F, sep=";")
```
