---
title: "Paired comparisons Wilcoxon"
author: "indra deCastro"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
      toc: true
      theme: united
---
      
<style type="text/css">
      .table {
            width: 100%;
      }
</style>
      
```{r setup, include=FALSE}
# setwd("/home/indra/Documents/20160101.Noriega.exp.remocion/")
knitr::opts_knit$set(root.dir ="/home/indra/Documents/20160101.Noriega.exp.remocion/", message=F, warning=F, verbose=F, comment=F)
knitr::opts_chunk$set(message=F, warning=F, verbose=F, echo=F, autodep=T, cache=T, fig.width=10)
options(markdown.HTML.header = system.file("misc", "datatables.html", package = "knitr"))
```

## Import data
```{r import, echo=T}
raw <- read.table("data/all.csv", header=T, sep=";")
```

## Select variables to compare
```{r, echo=F}
foc.names <- c("abun", "rich", "even",
               "rem.met2", # removal
               "altitud", # geographical
               "land.history", "cow.dens", "anthelmintic", # manegement
               "mpd.behav", "mpd.morph") 
```

Variables of interest (`r length(foc.names)`):

- taxonomic: abun, rich, even
- removal: rem.met2
- manegement: land.history, cow.dens, anthelmintic
- functional indices: mpd.behav, mpd.morph

## Natural log +1 transform all the variables of interest
```{r}
log.raw <- raw
log.raw[,foc.names] <- apply(log.raw[,foc.names], 2, function(x){log(x+1)})
```

## Calculate paired Wilcoxon test for 1 variable
(ie. abundance) to check the  
From the help-pages of `R`:  
*if both x and y are given and paired is FALSE, a Wilcoxon rank sum test (equivalent to the Mann-Whitney test: see the Note) is carried out.*  
**Thus, our analysis can be called Wilcoxon signed rank tests without any doubt.**
```{r, echo=T}
wilcox.test(log.raw$abun ~ log.raw$manejo, paired=T)
```

## Calculate paired Wilcoxon test for all variables
**ALL mean and sd (=standard deviation) values are given in ln(x+1)**
```{r}
all.wilcox <- sapply(foc.names, function(var){
      
      tmp <- complete.cases(log.raw[,c("manejo", var)])
      tmp <- log.raw[,c("site", "manejo", var)]
      tmp <- tmp[complete.cases(tmp),]
      tmp <- log.raw[rep(table(tmp$site)==2, each=2),c("manejo", var)]
      mywil <- wilcox.test(tmp[,var] ~ tmp$manejo, paired=T)
      summary.var <- list("mean.Int"=mean(tmp[tmp$manejo=="Int",var], na.rm=T),
                          "sd.Int"=sd(tmp[tmp$manejo=="Int",var], na.rm=T),
                          "mean.Ext"=mean(tmp[tmp$manejo=="Ext",var], na.rm=T),
                          "sd.Ext"=sd(tmp[tmp$manejo=="Ext",var], na.rm=T),
                          "V"=mywil$statistic, "pval"=round(mywil$p.value, 4))
      return(summary.var)
      })

wiltab <- t(all.wilcox)
wiltab <- data.frame("variable"=rownames(wiltab), wiltab, row.names=NULL)

# library(dplyr)
# ord.all.wilcox <- arrange(wiltab, pval)
# ord.all.wilcox
```


```{r, eval=T}
library(DT)
datatable(wiltab, width=35, options = list(pageLength = nrow(wiltab), dom="t")) %>%
      formatRound(columns=c("pval"), digits=4)
```


## Plots presenting raw values (unless log. is included in the name)
```{r, eval=T}
raw$log.abun <- log.raw$abun; raw$abun <- NULL
raw$log.cow.dens <- log.raw$cow.dens; raw$cow.dens <- NULL

foc.names <- c("log.abun", "rich", "even", "rem.met2", "altitud", "land.history", 
               "log.cow.dens", "anthelmintic", "mpd.behav", "mpd.morph")




numvars <- 1:length(foc.names)
library(reshape2)
long <- melt(raw[c("manejo", foc.names[numvars])])
library(dplyr)
long <- arrange(long, variable)
# both <- arrange(both, variable)
# 
# mylabs <- levels(long$variable)
# levels(long$variable) <- paste0(mylabs[numvars],
#                                " (w=", round(both$pval.wilcox[numvars], 2),
#                                ", t=", round(both$pval.ttest[numvars], 2), ")")

library(ggplot2)
ggplot(long, aes_(x=~manejo, y=~value)) +
      # geom_boxplot() +
      geom_violin() +
      # geom_line(aes(group = manejo)) + 
      facet_wrap(~variable, scales="free") +
      geom_dotplot(binaxis='y', stackdir='center', dotsize=0.5) +
      # geom_point() +
      labs(x="", y="", title="")

```

